---
title: "Tanley-Wood-Project2"
author: "Jordan Tanley and Jonathan Wood"
date: '2022-07-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here's the new file - idk what happened but my R malfunctioned for some reason and I had to make a new one. 

# Introduction - Jonathan



# Data - Jordan

```{r, include=FALSE}
library(readr)
library(tidyverse)
```

In order to read in the data using a relative path, be sure to have the data file saved in your working directory.  

```{r}
# read in the data
news <- read_csv("OnlineNewsPopularity/OnlineNewsPopularity.csv")
```

Next, let's subset the data so that we can only look at the data channel of interest.

```{r}
SocialMedia <- news %>% 
                as_tibble() %>% 
                filter(data_channel_is_socmed == 1) %>% 
                select(-c(url, timedelta, starts_with("data_channel")))
head(SocialMedia)
```


# Summarizations - Both



# Modeling



## Random Forest - Jordan




## Boosted Tree - Jonathan



# Comparison



# Automation






